---
title: "Toy full likelihood"
author: "Renata Diaz"
date: "8/6/2019"
output: github_document
---

```{r setup, include=FALSE}

library(MATSS)
library(matssldats)
library(LDATS)
source(here::here("analysis", "reports", "full_ldats_likelihood_stash", "full_ldats_likelihood.R"))

```

## Full model likelihood

Based on Slack & in person conversation with Hao Ye & Juniper Simonis.

Tacking a full likelihood computation on to existing runs of LDA_TS models, moving towards whole-model evaluation of goodness of fit (instead of evaluting and selecting LDA and TS components separately and sequentially). 

Likelihood of observed dataset (observed term frequencies in each document) from LDA and TS model...

```{r run multiple LDA and TS models, include = T, eval = F}
dat1 <- get_portal_rodents()

ldas <- LDA_set(dat1$abundance, topics = c(2, 5), nseeds = 1)

ts_fits <- run_TS(dat1, ldas, formulas = c("intercept", "time"), 
                  nchangepoints = c(0, 1), weighting = "proportional")

#plot(lda1)

#plot(ts1)
```

```{r load models, include = F, eval = T}
load(here::here("analysis", "reports", "full_ldats_likelihood_stash", "models.RData"))
```

```{r map ts fits to ldas}

ts_fit_names <- strsplit(names(ts_fits), split = ", ")

ts_fit_data <- data.frame(
    ts_index = 1:length(ts_fits),
    ts_names = unlist(names(ts_fits)),
    k = vapply(X = ts_fit_names, FUN = function(X) return(X[[1]]), FUN.VALUE = "k: 2"),
    seed = vapply(X = ts_fit_names, FUN = function(X) return(X[[2]]), 
                FUN.VALUE = "seed: 1"),
    formula = vapply(X = ts_fit_names, FUN = function(X) return(X[[3]]), 
                FUN.VALUE = "gamma ~ time"),
    changepoints = vapply(X = ts_fit_names, FUN = function(X) return(X[[4]]), 
                FUN.VALUE = "0 changepoints")
)

ts_fit_data$k <- substr(ts_fit_data$k, start = 4, stop = nchar(as.character(ts_fit_data$k)))
ts_fit_data$seed <- substr(ts_fit_data$seed, start = 7, stop = nchar(as.character(ts_fit_data$seed)))
ts_fit_data$changepoints <- substr(ts_fit_data$changepoints, start = 1, stop = nchar(as.character(ts_fit_data$changepoints))- 13)

ts_fit_data$lda_name <- paste0("k: ",
                               ts_fit_data$k, 
                               ", seed: ",
                               ts_fit_data$seed)
ts_fit_data$lda_index <- vapply(ts_fit_data$lda_name, function(x, ldamodels) return(which(names(ldamodels) == x)), ldamodels = ldas, FUN.VALUE = 1)

dat_to_aiccs <- function(data_row, ldamodels, tsmodels, datalist) {
    return(get_all_aiccs(lda_model = ldamodels[as.integer(data_row[8])],
                        ts_model = tsmodels[[as.integer(data_row[1])]],
                        counts_matrix = datalist$abundance))
}

```


```{r get aics}

model_aiccs <- apply(ts_fit_data, MARGIN = 1, FUN = dat_to_aiccs,
                     ldamodels = ldas, tsmodels = ts_fits, datalist = dat1)

model_aiccs <- as.data.frame(model_aiccs)

colnames(model_aiccs) <- ts_fit_data$ts_names

model_aiccs <- tidyr::gather(model_aiccs, key = "ts_names", value = "AICc") %>%
    dplyr::left_join(ts_fit_data, by = "ts_names")

```


```{r plot AICcs}

library(ggplot2)

all_aicc_plot <- ggplot(data = model_aiccs, aes(x = ts_index, y = AICc, color = k)) + 
    geom_point() +
    theme_bw()

all_aicc_plot

ts_index_summary <- model_aiccs %>%
    dplyr::select(ts_index, k, changepoints, formula) %>%
    dplyr::distinct()

ts_index_summary

```

```{r identify best model and close fits}

close_models <- model_aiccs %>%
    dplyr::mutate(AICc_close = 2 >= (AICc - min(AICc))) %>%
    dplyr::filter(AICc_close) %>%
    dplyr::mutate(AICc_ismin = AICc == min(AICc)) %>%
    dplyr::arrange(dplyr::desc(AICc_ismin))

close_models

```