---
title: "Toy full likelihood"
author: "Renata Diaz"
date: "8/6/2019"
output: github_document
---

```{r setup, include=FALSE}

library(MATSS)
library(drake)
library(matssldats)
library(LDATS)

db <- DBI::dbConnect(RSQLite::SQLite(), here::here("drake", "drake-cache.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)

```

## Full model likelihood

Based on Slack & in person conversation with Hao Ye & Juniper Simonis.

Tacking a full likelihood computation on to existing runs of LDA_TS models, moving towards whole-model evaluation of goodness of fit (instead of evaluting and selecting LDA and TS components separately and sequentially). 

Likelihood of observed dataset (observed term frequencies in each document) from LDA and TS model...

```{r pull in an LDA and a TS model}
dat1 <- readd(bbs_data_rtrg_1_11, cache = cache)
counts1 <- dat1$abundance
lda1 <- readd(lda_select_lda_bbs_data_rtrg_1_11_5, cache = cache)
ts1 <- readd(ts_bbs_data_rtrg_1_11_lda_select_lda_bbs_data_rtrg_1_11_5, cache = cache)[[4]]

#plot(lda1)

#plot(ts1)
```

```{r functions}

get_relevant_covars <- function(covariates_table, ts_formula) {
    formula_string <- unlist(as.character(ts_formula))
    
    formula_string <- formula_string[ which((formula_string != "~") & (formula_string != "gamma") & (formula_string != "1"))]

    covariates_table$intercept <- 1
    covars <- as.data.frame(covariates_table[ , c("intercept", formula_string)])
    
    return(covars)
}

get_theta <- function(ts_model, sim = 1){

covars <- get_relevant_covars(ts_model$data, ts_model$formula)

ncovar <- ncol(covars)
nseg <- ifelse(is.null(ts_model$rhos), 1, ncol(ts_model$rhos) + 1)
ntopics <- ncol(ts_model$data$gamma)

ndocs <- nrow(ts_model$data)

X <- matrix(nrow = ndocs, ncol = ncovar, data = unlist(covars), byrow = FALSE)

model_Eta <- ts_model$etas[sim, ]

Eta_matrix <- matrix(nrow = ncovar  * nseg, ncol = ntopics,
                     data = c(rep(0, times = ncovar * nseg), model_Eta), byrow = FALSE)

rho = ts_model$rhos[sim,]

tD <- unlist(ts_model$data[ , ts_model$timename])

Theta <- sim_TS_data(X, Eta_matrix, rho, tD, err = 0)

return(Theta) 
}


get_doc_lik <- function(counts_matrix, p_matrix, index) {
    counts <- as.integer(counts_matrix[index, ])
    ps <- p_matrix[index, ]
    doc_loglik <- dmultinom(x = counts, prob = ps, log = TRUE)
    return(doc_loglik)
}


get_aicc <- function(beta, theta, lda_model, ts_model) {
    P1exp <- theta %*% beta1exp

doc_lik1 <- vapply(1:nrow(counts1), FUN = get_doc_lik, counts_matrix = counts1, 
                   p_matrix = P1exp, FUN.VALUE = -1)

ldapars <- attr(logLik(lda_model[[1]]), "df")
tspars <- attr(logLik(ts_model), "df")

totalpars <- ldapars + tspars

nobs <- attr(logLik(ts_model), "nobs")

AICc <- (-2 * (sum(doc_lik1))) + 2*(totalpars) + 
    (2 * totalpars^2 + 2 * totalpars)/(nobs - totalpars - 1)

return(AICc)
}

```

```{r get beta}
beta1 <- lda1$`k: 5, seed: 122`@beta

beta1exp <- exp(beta1)
```


```{r get all thetas}
all_thetas <- lapply(X = c(1:nrow(ts1$etas)), FUN = get_theta, ts_model = ts1)
```

```{r get AICcs}
all_aicc <- vapply(X = all_thetas, FUN = get_aicc, beta = beta1exp, lda_model = lda1, ts_model = ts1, FUN.VALUE = 100.1)

```

```{r plot AICcs}

library(ggplot2)

all_aicc <- as.data.frame(all_aicc)

aicc_plot <- ggplot(data = all_aicc, aes(x = all_aicc)) + 
    geom_histogram() +
    theme_bw()

all_aicc

```


```{r get all model aiccs}
ts_all <- readd(ts_bbs_data_rtrg_1_11_lda_select_lda_bbs_data_rtrg_1_11_5, cache = cache)

get_all_aiccs <- function(lda_model, ts_model, dat_list) {
    betaexp <- exp(lda_model[[1]]@beta)
    all_thetas <- all_thetas <- lapply(X = c(1:nrow(ts_model$etas)), FUN = get_theta, ts_model = ts_model)
    all_aicc <- vapply(X = all_thetas, FUN = get_aicc, beta = betaexp, lda_model = lda_model, ts_model = ts_model, FUN.VALUE = 100.1)
    return(all_aicc)
}

all_aicc <- get_all_aiccs(lda1, ts1, dat1)

all_model_aiccs <- lapply(ts_all, FUN = get_all_aiccs, lda_model = lda1, dat_list = dat1)

all_model_aiccs1 <- dplyr::bind_rows(all_model_aiccs) %>%
    tidyr::gather(key = "model", value = "AICc")

all_aicc_plot <- ggplot(data = all_model_aiccs1, aes(x = model, y = AICc, color = model)) + 
    geom_point() +
    theme_bw()

all_aicc_plot

```